---
title: "2025-07-25_reviewer_changes"
author: "Lucore"
date: "2024-07-25"
output: html_document
---

# Notes 

Up to date code for neopterin manuscriptpost reviews from Science Advances. 

# Load environment

```{r}

#libraries
library(tidyverse)
library(lme4)
library(bbmle)
library(mgcv)
library(broom.mixed)
library(cowplot)
library(sjPlot)
library(gratia)
library(ggpubr)
library(MuMIn)
library(ggiraph)
library(flextable)


n <- read_csv("data/2024-07-25_compiled_neopterin.csv")

m <- read_csv("data/MERRA2_maxt.csv")

# functions
source("functions/my_scale.R")

source("functions/extract_model_info.R")

source("functions/plot_theme.R")

source("functions/temp_bar.R")

```

# 1. Prep data

## 1a. remove outliers

```{r}

# Define mean and standard deviation
mean <- mean(n$neo_nmol_SG)

std_dev <- sd(n$neo_nmol_SG)  

# Calculate 2 standard deviations above the mean
value_2std_above_mean <- mean + (4 * std_dev)

# Filter samples
n1 <- n %>%
  filter(neo_nmol_SG <= value_2std_above_mean) %>%
  filter(ID != "TET")

```


# 2. Choose best fitting temp time interval 

Min and max temp show the same trends, average temp 15 days before sample collection is the best fitting time interval. Min and Max temperature show the same shape. 

## 2a. Prep data

```{r}

n_temp <- n1 %>%
  filter(!is.na(max_t_days1)) %>% # remove days whens theres no temp observation, only for model compairson purposes
  mutate(across(c(max_t_days1:rain_days4, age), my_scale, .names = "{.col}_s"),
         ID = as.factor(ID),
         group = as.factor(group))

```

## 2b. Evaluate best max temp interval 

Max temp averages 15days before sample collection is the best fitting time interval. 

```{r}

# List of variables to iterate over
variables <- c("max_t_days1_s", "max_t_days10_s", "max_t_days11_s", "max_t_days12_s", "max_t_days13_s", "max_t_days14_s", "max_t_days15_s", "max_t_days16_s", "max_t_days17_s", "max_t_days18_s", "max_t_days19_s", "max_t_days2_s", "max_t_days20_s", "max_t_days21_s", "max_t_days28_s", "max_t_days3_s", "max_t_days35_s", "max_t_days5_s",  "max_t_days7_s", "max_t_days9_s")

# Initialize a list to store the models
models <- list()
plots <- list()

# Loop through each variable, build the model, and store it in the list
for (var in variables) {
  model_name <- paste0("g_", gsub("_s", "", var))
  formula <- as.formula(paste("neo_nmol_SG ~ s(age_s) + group + s(", var, ") + s(ID, bs = 're')"))
  models[[model_name]] <- gam(formula, data = n_temp, method = "REML", family = Gamma(link = "log"))
  
  plots[[model_name]] <- plot(models[[model_name]], select = 2, rug = FALSE)
}

# Extract model names
model_names <- names(models)

# Print summary of each model
lapply(models, summary)

# Create AIC table
max_aic_results <- AICctab(models, weights = TRUE)

# Print AIC table
print(max_aic_results)  

# Print model plots
for (model_name in model_names) {
  print(plots[[model_name]])
}

```


## 2c. Evaluate best min temp interval  

Min temp averages 15days before sample collection is the best fitting time interval.

```{r}

# List of variables to iterate over
variables <- c("min_t_days1_s", "min_t_days10_s", "min_t_days11_s", "min_t_days12_s", "min_t_days13_s", "min_t_days14_s", "min_t_days15_s", "min_t_days16_s", "min_t_days17_s", "min_t_days18_s", "min_t_days19_s", "min_t_days2_s", "min_t_days20_s", "min_t_days21_s", "min_t_days28_s", "min_t_days3_s", "min_t_days35_s", "min_t_days5_s",  "min_t_days7_s", "min_t_days9_s")

# Initialize a list to store the models
models <- list()
plots <- list()

# Loop through each variable, build the model, and store it in the list
for (var in variables) {
  model_name <- paste0("g_", gsub("_s", "", var))
  formula <- as.formula(paste("neo_nmol_SG ~ s(age_s) + group + s(", var, ") + s(ID, bs = 're')"))
  models[[model_name]] <- gam(formula, data = n_temp, method = "REML", family = Gamma(link = "log"))
  
  plots[[model_name]] <- plot(models[[model_name]], select = 2, rug = FALSE)
  
}

# Extract model names
model_names <- names(models)

# Print summary of each model
lapply(models, summary)

# Create AIC table
min_aic_results <- AICctab(models, weights = TRUE)

# Print AIC table
print(min_aic_results) 

# Print model plots
for (model_name in model_names) {
  print(plots[[model_name]])
}

```

## 2c. Compare results for min and max temp 
### 2c i. Build models outside of loop

Creates figure S1

```{r}

g_min_15 <- gam(neo_nmol_SG ~ s(age_s) + group + s(min_t_days15_s) + s(ID, bs = "re"),
             data = n_temp,
             method = "REML",
             family = Gamma(link = "log"))
min_t_15 <- summary(g_min_15)
min_t_15 

min_plot15 <- draw(g_min_15, select = 2, rug= FALSE)
min_plot <- min_plot15 +
  labs(
    x = "Average min temp scaled (°C)",
    y = "Neopterin (ng/ml) partial effects") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black"),
        axis.title.y = element_text(size = 25, face = "bold"),
        axis.title.x = element_text(size = 25, face = "bold"),
        axis.text = element_text(size = 23, color = "black"),
        plot.title = element_blank()) 

g_max_15 <- gam(neo_nmol_SG ~ s(age_s) + group + s(max_t_days15_s) + s(ID, bs = "re"),
             data = n_temp,
             method = "REML",
             family = Gamma(link = "log"))
max_t_15 <- summary(g_max_15)
max_t_15 

max_plot15 <- plot(g_max_15, 
     select = 2, #selections 2nd smooth function
     rug= FALSE)

max_plot15 <- draw(g_max_15, select = 2, rug= FALSE)
max_plot <- max_plot15 +
  labs(
    x = "Average max temp scaled (°C)",
    y = " ") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black"),
        axis.title.y = element_text(size = 25, face = "bold"),
        axis.title.x = element_text(size = 25, face = "bold"),
        axis.text = element_text(size = 23, color = "black"),
        plot.title = element_blank()) 

# combine plots side by side

combined_plot <- ggarrange(min_plot, max_plot, ncol = 2)

# save figure S1

tiff(file= "figures/figure_S1", width=15, height=9, units="in", res=500)  

combined_plot 

dev.off()



```

### 2c ii. Compare AICc tables

Creates table S4

```{r}

min_aic <- min_aic_results %>% # Create AIC tables 
  as.data.frame() %>%
  rownames_to_column() %>%
  rename(min_t_model = rowname,
         min_t_dAICc = dAICc,
         min_t_weight = weight) %>%
  mutate(min_t_model = str_replace(min_t_model, ".*_(days\\d+)", "\\1"),
         min_t_dAICc = round(min_t_dAICc, 2),
         min_t_weight = round(min_t_weight, 2)) %>%
  select(-df)

max_aic <- max_aic_results %>%
  as.data.frame() %>%
  rownames_to_column() %>%
  rename(max_t_model = rowname,
         max_t_dAICc = dAICc,
         max_t_weight = weight) %>%
  mutate(max_t_model = str_replace(max_t_model, ".*_(days\\d+)", "\\1"),
         max_t_dAICc = round(max_t_dAICc, 2),
         max_t_weight = round(max_t_weight, 2)) %>%
  select(-df)

# Make flex table
aic <- min_aic %>% # bind aic tables
  cbind(max_aic) %>%
  flextable() %>%
  vline(j = 3) %>%
  bold(i = 1, part = "header")

print(aic, preview = "docx")  


```


# 3. Neopterin: Generalized Additive Model (GAM)

## 3a. Prep data

```{r}

# 1b scale variables # this needs to be changed for other parts of analysis 

n2 <- n1 %>%
  select(c(sample:repro, max_t_days15, min_t_days15, rain_days4)) %>%
  mutate(across(c(max_t_days15:rain_days4, age), my_scale, .names = "{.col}_s"),
         ID = as.factor(ID),
         group = as.factor(group))
  
```


## 3b. Build model

```{r}

m_gam <- gam(neo_nmol_SG ~ s(age_s) + group + s(max_t_days15_s) + s(ID, bs = "re"),
             data = n2,
             method = "REML",
             family = Gamma(link = "log"))

summary(m_gam)

```


## 3c. Plot: Figure 1

```{r}

plot1 <- draw(m_gam, 
     select = 2, #selections 2nd smooth function
     rug= FALSE) 

glmm_1_temp <- plot1 + 
  ggtitle(" ") +
  xlab("Average max temp (°C)") +
  ylab("Partial effects of neopterin (ng/mL)") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black", linewidth = .28),
        axis.title.y = element_text(size = 8),
        axis.title.x = element_text(size = 8),
        axis.text = element_text(size = 8, color = "black"),
        text = element_text(family = "sans")) +
  geom_vline(xintercept = 0.7523, color = "#a6000e", linetype = "dashed", linewidth = .5) +
  scale_x_continuous(breaks = c(-0.6866, -0.3286, 0.0344, 0.3862, 0.7523, 1.13308), 
                     labels = c(26, 27, 28, 29, 30, 31))

ggsave("figures/figure_1", plot = glmm_1_temp, device = "pdf", width=9, height=8, units = "cm", dpi = 1200, colormodel = "rgb")


```


# 4. Neopterin: Generalized Linear Mixed Models (GLMM's)

## 4a. Build and compare models

Added reproductive status, and rain into models. Rain controls for seasonality. Repro status also controls for sex because there is a Male variable

```{r}

glmm_0 <- glmer(neo_nmol_SG ~ group + poly(age_s, degree = 2) + rain_days4_s + repro + (1|ID),
                data = n2,
                family = Gamma(link = "log"))
summary(glmm_0)

glmm_1 <- glmer(neo_nmol_SG ~ group + poly(age_s, degree = 2) + rain_days4_s + repro + max_t_days15_s + (1|ID),
                data = n2,
                family = Gamma(link = "log"))
summary(glmm_1)

glmm_2 <- glmer(neo_nmol_SG ~ group + poly(age_s, degree = 2) * max_t_days15_s + rain_days4_s + repro + (1|ID),
                data = n2,
                family = Gamma(link = "log"))
summary(glmm_2)

AICctab(glmm_0, glmm_1, glmm_2, weights = TRUE)

```

## 4b. Average best fitting models

```{r}

glmm_avg <- model.avg(glmm_1, glmm_2)
  
glmm_avg <- as.data.frame(coef(glmm_avg ))

AICctab(glmm_1, glmm_2, weights = TRUE)

```

## 4c. Plot: Figure 2

```{r}

# Assemble data for coef plot 
# glmm_1

merge_1 <- extract_model_info(glmm_1, conf_level_1 = 0.5, conf_level_2 = 0.95) %>%
  mutate(model = "glmm_1")

# glmm_2

merge_2 <- extract_model_info(glmm_2, conf_level_1 = 0.5, conf_level_2 = 0.95) %>%
  mutate(model = "glmm_2")

# join model info and filter variables 

models <- rbind(merge_1, merge_2) %>%
  filter(term %in% c("poly(age_s, degree = 2)2", "max_t_days15_s", "poly(age_s, degree = 2)2:max_t_days15_s", 	
"rain_days4_s"))

# plot the models 

coef_plot <- ggplot(models, aes(x = term, y = estimate, color = model)) +
  geom_hline(yintercept = 0, linetype = "dashed", size = .5, color = "grey") +
  geom_point(position = position_dodge(width = 0.3), size = 1) +
  geom_linerange(aes(ymin = conf95.low, ymax = conf95.high),
                 position = position_dodge(width = 0.3), linewidth = .35) +
  geom_linerange(aes(ymin = conf50.low, ymax = conf50.high),
                 position = position_dodge(width = 0.3), linewidth = .75) + 
  scale_color_manual(values = c("#1b7837", "#762a83")) +
  ylab("Estimate") +
  coord_flip() +
  theme_minimal() +
  theme(axis.line = element_line(color = "black", linewidth = .28),
              axis.text = element_text(size = 8, color = "black"), 
              text = element_text(size = 8, color = "black", family = "sans"), 
              panel.grid.minor = element_blank(),
              panel.grid.major = element_blank(),
              legend.position = "none",
              axis.title.y = element_blank()) +
  scale_x_discrete(breaks = c('poly(age_s, degree = 2)2:max_t_days15_s',  
                              'poly(age_s, degree = 2)2', 
                              "max_t_days15_s",
                              "rain_days4_s"), 
                    labels = c(expression("Age" ^2 * " * max temp"), 
                               expression("Age"^2), 
                               "Max temp",
                               "Rain")) 
                   
# add custom legend

coef_plot <- coef_plot +
  annotate(geom = "point", x = 4, y = 2, shape = 16, size = 1, color = "#762a83") +
  annotate(geom = "segment", x = 4, xend = 4, y = 1.7, yend = 2.3, size = .3, color = "#762a83") +
  annotate(geom = "text", x = 4, y = 4.9, label = bquote("m2: " * Delta * "AIC = 0.0 " * omega * " =  0.975"), size = 1.7, family = "sans") +
  
  annotate(geom = "point", x = 3.7, y = 2, shape = 16, size = 1, color = "#1b7837") +
  annotate(geom = "segment", x = 3.7, xend = 3.7, y = 1.7, yend = 2.3, size = .3, color = "#1b7837") +
  annotate(geom = "text", x = 3.7, y = 4.9, label = bquote("m1: " * Delta * "AIC = 7.3 " * omega * " =  0.025"), size = 1.7, family = "sans") 
  
ggsave("figures/figure_2", plot = coef_plot, device = "pdf", width=9, height=5, units = "cm", dpi = 1200, colormodel = "rgb")


```


## 4d. Plot: Figure 3

### 4d i. Prep data

```{r}

# Add predictions from best model
n2$preds <- predict(glmm_2, type = "response")

# create temp bins
bin1 <- n2 %>% filter(max_t_days15 >= 25.5 & max_t_days15 <= 27)

bin2 <- n2 %>% filter(max_t_days15 >= 27 & max_t_days15 <= 28.5)

bin3 <- n2 %>% filter(max_t_days15 >= 28.5 & max_t_days15 <= 30)

bin4 <- n2 %>% filter(max_t_days15 >= 30 & max_t_days15 <= 31.5)

```

### 4d ii. Make plots

```{r}

# plot each bin 
plot_1 <- ggplot(bin1, aes(x = age, y = preds)) +
  geom_point(color = "#e7ba8c", alpha = 0.4, size = 1) +
  geom_smooth(aes(y = preds), method='lm', formula = y~poly(x,2), color = "#e7ba8c", linewidth = .5) +
  ylim(0, 2300) +
  scale_x_continuous(breaks = c(0, 10, 20), labels = c("0", "10", "20")) +
  coord_cartesian(xlim = c(0, 25)) +
  ylab("Neopterin (ng/mL)") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black", linewidth = 0.28),
        axis.title.x = element_blank(),
        axis.text = element_text(size = 8, color = "black"),
        axis.ticks = element_line(color = "black", linewidth = 0.28),
        axis.ticks.length.y = unit(.1, "cm")) +
  annotate("text", x = 3, y = 2300, label = "A", color = "black", size = 5, fontface = "bold", family = "sans")

plot_2 <- ggplot(bin2, aes(x = age, y = preds)) +
  geom_point(color = "#e9966a", alpha = 0.4, size = 1) +
  geom_smooth(aes(y = preds), method='lm', formula = y~poly(x,2), color = "#e9966a", linewidth = .5) +
  ylim(0, 2300) +
  scale_x_continuous(breaks = c(0, 10, 20), labels = c("0", "10", "20")) +
  coord_cartesian(xlim = c(0, 25)) +
  plot_theme() +
  annotate("text", x = 3, y = 2300, label = "B", color = "black", size = 5, fontface = "bold") 

plot_3 <- ggplot(bin3, aes(x = age, y = preds)) +
  geom_point(color = "#d9735a", alpha = 0.4, size = 1) +
  geom_smooth(aes(y = preds), method='lm', formula = y~poly(x,2), color = "#d9735a", linewidth = .5) +
  ylim(0, 2300) +
  scale_x_continuous(breaks = c(0, 10, 20), labels = c("0", "10", "20")) +
  coord_cartesian(xlim = c(0, 25)) +
  plot_theme() +
  annotate("text", x = 3, y = 2300, label = "C", color = "black", size = 5, fontface = "bold") 

plot_4 <- ggplot(bin4, aes(x = age, y = preds)) +
  geom_point(color = "#a6000e", alpha = 0.4, size = 1) +
  geom_smooth(aes(y = preds), method='lm', formula = y~poly(x,2), color = "#a6000e", linewidth = .5) +
  ylim(0, 2300) +
  scale_x_continuous(breaks = c(0, 10, 20), labels = c("0", "10", "20")) +
  coord_cartesian(xlim = c(0, 25)) +
  plot_theme() +
  annotate("text", x = 3, y = 2300, label = "D", color = "black", size = 5, fontface = "bold") 

# plot overlapping trends

plot_5 <- ggplot(bin4, aes(x = age, y = preds)) +
  ylim(0, 2300) +
  scale_x_continuous(breaks = c(0, 10, 20), labels = c("0", "10", "20")) +
  coord_cartesian(xlim = c(0, 25)) +
  geom_smooth(data = bin1, aes(y = preds), method='lm', formula = y~poly(x,2), color = "#e7ba8c", se = FALSE, linewidth = .5) +
  geom_smooth(data = bin2, aes(y = preds), method='lm', formula = y~poly(x,2), color = "#e9966a", se = FALSE, linewidth = .5) +
  geom_smooth(data = bin3, aes(y = preds), method='lm', formula = y~poly(x,2), color = "#d9735a", se = FALSE, linewidth = .5) +
  geom_smooth(aes(y = preds), method='lm', formula = y~poly(x,2), color = "#a6000e", se = FALSE, linewidth = .5) +
  plot_theme() +
  annotate("text", x = 3, y = 2300, label = "E", color = "black", size = 5, fontface = "bold", family = "sans") 

```

### 4d iii. Make bars

```{r}

data <- data.frame(start = 1, end = 21.5)

bar_1 <- temp_bar(data, "#e7ba8c", "25.5 - 27°C") +
  theme(text = element_text(size = 8)) 
  
bar_2 <- temp_bar(data, "#e9966a", "27 - 28.5°C") 

bar_3 <- temp_bar(data, "#d9735a", "28.5 - 30°C") 

bar_4 <- temp_bar(data, "#a6000e", "30 - 31.5°C") 

bar_5 <- temp_bar(data, "white", "")

```

### 4d iv. Combine plots 

```{r}

panel_1 <- ggarrange(bar_1, plot_1, nrow = 2, ncol = 1,
                             heights = c(1,8),
                 align = "v")

panel_2 <- ggarrange(bar_2, plot_2, nrow = 2, ncol = 1,
                             heights = c(1,8))

panel_3 <- ggarrange(bar_3, plot_3, nrow = 2, ncol = 1,
                             heights = c(1,8))

panel_4 <- ggarrange(bar_4, plot_4, nrow = 2, ncol = 1,
                             heights = c(1,8))

panel_5 <- ggarrange(bar_5, plot_5, nrow = 2, ncol = 1,
                             heights = c(1,8),
                 align = "v")

#combine plots 1-4
p1 <- ggarrange(panel_1, panel_2, panel_3, panel_4, nrow = 1, ncol = 4,
                align = "h",
          widths = c(1, .73, .73, .73))

fig <- ggarrange(p1, panel_5, nrow = 1, ncol = 2,
          align = "h",
          widths = c(1, 0.3)) +
          theme(plot.margin = margin(0.1,0.1,0.8,0.1, "cm"))  


fig_final <- annotate_figure(fig, fig.lab = "Age (years)",
                fig.lab.size = 10,
                fig.lab.pos = "bottom") +
    theme(plot.margin = margin(0.1,0.1,0.1,0.1, "cm"))



ggsave("figures/figure_3", plot = fig_final, device = "pdf", width=18.4, height=9.2, units = "cm", dpi = 1200, colormodel = "rgb")


```


# 5. Climate data: Linear mixed models (LMM's) 

## 5a. Standardize climate data

```{r}

# clean MERRA data 
m1 <- m %>%
  select(c(Temperature, Date)) %>%
  rename(temp = Temperature,
         date = Date) %>%
  mutate(temp = temp - 273.15,
         date = ymd(date),
         month = month(date),
         year = year(date),
         month_year = as.Date(make_date(year, month)))

# calculate deviation from PV data 
mean <- mean(n2$max_t_days15) # get mean of palo verde data

m2 <- m1 %>%
  mutate(difference = temp - mean) %>% #find distance from mean
  mutate(square = difference ^2)  # square distance

sum <- sum(m2$square) # sum the distances

deviation <- sum/16103 # find deviation from mean of palo verde data 

m3 <- m2 %>%
  mutate(adj_temp = temp + deviation)

# Count of days over 30 C
count_1980_2014 <- m3 %>%
  filter(year < 2014) %>%
  filter(adj_temp >= 30) %>%
  summarize(total_rows = n()) # 2311 days 

2311/12419 # 18 percent of days above 30 C

count_2014_2024 <- m3 %>%
  filter(year > 2014) %>%
  filter(adj_temp >= 30) %>%
  summarize(total_rows = n()) # 1127 days 

1127/3319 # 34% of days above 30 C in past decade 

```

## 5b. Build models

```{r}

m_intercept <- lm(adj_temp ~ 1, data = m3)

m_date <- lm(adj_temp ~ date, data = m3)
summary(m_date)

AICtab(m_intercept, m_date, weights = TRUE)

#extract prediction
pred_ci <- predict(m_date, interval = "confidence", level = 0.95)

# Add predictions and CI 
m3$preds <- pred_ci[, "fit"]
m3$ci_lower <- pred_ci[, "lwr"]
m3$ci_upper <- pred_ci[, "upr"]

```

## 5c. Plot: Figure 4

```{r}

# adjust data to plot temp observation by month instead of day
m_month <- m3 %>%
  group_by(year, month) %>%
  summarise(mean_month = mean(adj_temp)) %>%
  mutate(month_year2 = make_date(year, month))

label_colors <- c("black", "black", "black", "#a6000e", "black")

plot <- ggplot(m_month, aes(month_year2, mean_month)) +
  geom_ribbon(aes(ymax = ifelse(mean_month > 30, mean_month, 30), ymin = 30), fill = "#a6000e") +
  geom_ribbon(aes(ymax = ifelse(mean_month < 30, mean_month, 30), ymin = -Inf), fill = "#E9EAEC") +
  geom_line(linewidth = .28) +
  geom_hline(yintercept = 30, linetype = "dashed", color = "#a6000e", linewidth = .28) +
  geom_smooth(data = m3, aes(x = month_year, y = preds), color = "black", linewidth = .4) +
  ylab("Avg max temperature per month (°C)") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.title.x = element_blank(),
        text = element_text(size = 8, family = "sans"),
        axis.line = element_line(color = "black", linewidth = .28),
        axis.text.x = element_text(size = 8, color = "black"),
        axis.text.y = element_text(size = 8, color = label_colors)) +
  annotate("text", x = as.Date("2015-01-01"), y = 26.8, label = bquote(beta == 5.05e-05), size = 2.5, color = "black", family = "sans") +
  annotate("text", x = as.Date("2015-01-01"), y = 26.5, label = "SE = 1.66e-06", size = 2.5, color = "black", family = "sans") 

ggsave("figures/figure_4", plot = plot, device = "pdf", width=14.4, height=8.2, units = "cm", dpi = 1200, colormodel = "rgb")

```


# 6. Extended data

## 6a. Plot: Extended data figure 2

Coefficient plot including all model variables. 

```{r}

models_full <- rbind(merge_1, merge_2) %>%
  filter(term != "(Intercept)")
 
# plot the models 

coef_plot_full <- ggplot(models_full, aes(x = term, y = estimate, color = model)) +
  geom_hline(yintercept = 0, linetype = "dashed", size = 1, color = "grey") +
  geom_point(position = position_dodge(width = 0.3), size = 3) +
  geom_linerange(aes(ymin = conf95.low, ymax = conf95.high),
                 position = position_dodge(width = 0.3), linewidth = 1) +
  geom_linerange(aes(ymin = conf50.low, ymax = conf50.high),
                 position = position_dodge(width = 0.3), linewidth = 2) + 
  scale_color_manual(values = c("#1b7837", "#762a83")) +
  ylab("Estimate") +
  coord_flip() +
  theme_minimal() +
  theme(axis.line = element_line(color = "black"),
              axis.text = element_text(size = 23, color = "black"), 
              text = element_text(size = 23, color = "black"), 
              panel.grid.minor = element_blank(),
              panel.grid.major = element_blank(),
              legend.position = "none",
              axis.title.y = element_blank()) +
  scale_x_discrete(breaks = c('reproPregnant',
                              'reproM',
                              'reproLactating',
                              'reprojuv',
                              'rain_days4_s',
                              'poly(age_s, degree = 2)2:max_t_days15_s',  
                              'poly(age_s, degree = 2)2', 
                              'poly(age_s, degree = 2)1:max_t_days15_s',
                              'poly(age_s, degree = 2)1',
                              'max_t_days15_s',
                              'groupTenori',
                              'groupPalmas'),
                    labels = c("pregnant",
                               "sex:male",
                               "lactating",
                               "juvenile female",
                               "rain",
                               expression("Age" ^2 * " * max temp"), 
                               expression("Age"^2), 
                               "Age * max temp",
                               "Age",
                               "Max temp",
                               "group: Tenori",
                               "group: Palmas",
                               "rain"))
                               
                               # add custom legend

coef_plot_full <- coef_plot_full +
  annotate(geom = "point", x = 2.2, y = -6, shape = 16, size = 5, color = "#762a83") +
  annotate(geom = "segment", x = 2.2, xend = 2.2, y = -5.7, yend = -6.3, size = 3, color = "#762a83") +
  annotate(geom = "text", x = 2.2, y = -3.2, label = "m2: ΔAIC = 0.0, ω =  0.975 ", size = 5) +
  
  annotate(geom = "point", x = 1.8, y = -6, shape = 16, size = 5, color = "#1b7837") +
  annotate(geom = "segment", x = 1.8, xend = 1.8, y = -5.7, yend = -6.3, size = 3, color = "#1b7837") +
  annotate(geom = "text", x = 1.8, y = -3.2, label = "m1: ΔAIC = 7.3, ω =  0.025", size = 5) 


tiff(file="figures/figure_S2", width=10, height=8, units="in", res=500)


coef_plot_full


dev.off()


```


## 6b. Analysis with oldest individual included

### 6b i. GAM results 

```{r}

# prep full data 

n_full <- n %>%
  filter(neo_nmol_SG <= value_2std_above_mean) %>%
  mutate(across(c(max_t_days1:rain_days4, age), my_scale, .names = "{.col}_s"),
         ID = as.factor(ID),
         group = as.factor(group))

# create model

m_gam_full <- gam(neo_nmol_SG ~ s(age_s) + group + s(max_t_days15_s) + s(ID, bs = "re"),
             data = n_full,
             method = "REML",
             family = Gamma(link = "log"))

# compare model results 

variables <- c("s(age_s)", "s(max_t_days15_s)", "s(ID)")

full <- tidy(m_gam_full) %>%
  select(edf, ref.df, statistic) %>%
  mutate(model = "gam_full") %>%
  cbind(variables) %>%
  mutate('R-sq' = 0.395) %>%
  select(variables, edf, ref.df, statistic, model, 'R-sq') %>%
  filter(variables != "s(ID)") %>%
  mutate(variables = case_when(variables == 's(age_s)' ~ 'spline: age',
                          variables == 's(max_t_days15_s)' ~ 'spline: max temp')) 

gam_comparison <- tidy(m_gam) %>%
  select(edf, ref.df, statistic) %>%
  mutate(model = "gam_filtered") %>%
  cbind(variables) %>%
  mutate('R-sq' = 0.364) %>%
  select(model, variables, edf, ref.df, statistic, 'R-sq') %>%
  filter(variables != "s(ID)") %>%
  mutate(variables = case_when(variables == 's(age_s)' ~ 'spline: age',
                          variables == 's(max_t_days15_s)' ~ 'spline: max temp')) %>%
  rbind(full) %>%
  mutate(edf = round(edf, 2),
         ref.df = round(ref.df, 2),
         statistic = round(statistic, 2))
 

# make table

gam_comparison_table <- gam_comparison  %>%
  flextable() %>%
  hline(i = 2) %>%
  vline(j = 1, part = "body") %>%
  merge_v(j = c(1, 6)) %>%
  fix_border_issues() %>%
  bold(i = 1, part = "header")

print(gam_comparison_table, preview = "docx")

```

### 6b ii GLMM results

```{r}

# rebuild models with full dataset
ctrl <- glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1750)) # up iterations for model to converge

glmm_0_full <- glmer(neo_nmol_SG ~ group + poly(age_s, degree = 2) + rain_days4_s + repro + (1|ID),
                data = n_full,
                family = Gamma(link = "log"))

glmm_1_full <- glmer(neo_nmol_SG ~ group + poly(age_s, degree = 2) + rain_days4_s + max_t_days15_s + (1|ID),
                data = n_full,
                family = Gamma(link = "log"),
                 control = ctrl)

glmm_2_full <- glmer(neo_nmol_SG ~ group + poly(age_s, degree = 2) * max_t_days15_s + rain_days4_s + (1|ID),
                data = n_full,
                family = Gamma(link = "log"),
                 control = ctrl)

AICctab(glmm_0_full, glmm_1_full, glmm_2_full, weights = TRUE)

# compare AICc

aic_filtered <- as.data.frame(AICctab(glmm_1, glmm_2, weights = TRUE)) %>%
  rownames_to_column("model") %>%
  mutate(weight = weight * 100) %>%
  mutate(weight = round(weight, 2)) %>%
  mutate(dAICc = round(dAICc, 2)) %>%
   mutate(model = case_when(model == 'glmm_1' ~ 'glmm_1_filtered',
                          model == 'glmm_2' ~ 'glmm_2_filtered'))

aic_comparison <- as.data.frame(AICctab(glmm_1_full, glmm_2_full, weights = TRUE)) %>%
  rownames_to_column("model") %>%
  mutate(weight = weight * 100) %>%
  mutate(weight = round(weight, 2)) %>%
  mutate(dAICc = round(dAICc, 2)) %>%
  rbind(aic_filtered)

# make table

aic_comparison_table <- aic_comparison %>%
  flextable() %>%
  hline(i = 2) %>%
  vline(j = 1, part = "body") %>%
  bold(i = 1, part = "header")

print(aic_comparison_table, preview = "docx")


# compare main results of best fitting models

glmm_2_full_summary <- tidy(glmm_2_full) %>%
  select(term, estimate, std.error) %>%
   mutate(model = "glmm_2_full")

glmm_2_filtered_summary <- tidy(glmm_2) %>%
  select(term, estimate, std.error) %>%
   mutate(model = "glmm_2_filered") %>% 
  filter(term != "reprojuv" & term != "reproLactating" & term != "reproM" & term != "reproPregnant")


glmm_2_comparison <- rbind(glmm_2_full_summary, glmm_2_filtered_summary) %>%
  filter(!is.na(std.error)) %>% # remove intercept and random effect rows 
  filter(term %in% c('max_t_days15_s', 'poly(age_s, degree = 2)2', 'poly(age_s, degree = 2)2:max_t_days15_s', "rain_days4_s")) %>%
  mutate(term = case_when(term == 'rain_days4_s' ~ "Rain",
                          term == 'max_t_days15_s' ~ 'max temp',
                          term == 'poly(age_s, degree = 2)2' ~ "Age squared",
                          term == 'poly(age_s, degree = 2)2:max_t_days15_s' ~ "Age squared * max temp")) %>%
  mutate(estimate = round(estimate, 2),
         std.error = round(std.error, 2)) %>%
  rename(predictor = term) %>%
  select(model, predictor, estimate, std.error)


# make table 

glmm_2_comparison_table <- glmm_2_comparison %>%
  flextable() %>%
  hline(i = 4) %>%
  vline(j = 1, part = "body") %>%
  bold(i = 1, part = "header")

print(glmm_2_comparison_table, preview = "docx")

```


## 6c. Extended data table 4

Results from all three GLMM's 

```{r}

# make dataframe with AIC results 

aic <- as.data.frame(AICctab(glmm_0, glmm_1, glmm_2, weights = TRUE)) %>%
  rownames_to_column("model") %>%
  mutate(weight = weight * 100) %>%
  mutate(weight = round(weight, 2))

# Make dataframe with model results

m0_summary <- tidy(glmm_0) %>%
  select(term, estimate, std.error) %>%
   mutate(model = "glmm_0")

m1_summary <- tidy(glmm_1) %>%
  select(term, estimate, std.error) %>%
   mutate(model = "glmm_1")

m2_summary <- tidy(glmm_2) %>%
  select(term, estimate, std.error) %>%
   mutate(model = "glmm_2")


m_summary <- rbind(m1_summary, m2_summary) %>%
  filter(!is.na(std.error)) %>% # remove intercept and random effect rows 
  filter(term != "(Intercept)") %>%
  mutate(term = case_when(term == 'groupPalmas' ~ 'group: Palmas',
                          term == 'groupTenori' ~ 'group: Tenori',
                          term == 'poly(age_s, degree = 2)1' ~ 'Age',
                          term == 'max_t_days15_s' ~ 'max temp',
                          term == 'poly(age_s, degree = 2)2' ~ "Age squared",
                          term == 'poly(age_s, degree = 2)1:max_t_days15_s' ~ "Age * max temp",
                          term == 'poly(age_s, degree = 2)2:max_t_days15_s' ~ "Age squared * max temp",
                          term == 'rain_days4_s' ~ "Rain",
                          term == 'reprojuv' ~ 'Juvenile female',
                          term == 'reproLactating' ~ 'Lactating',
                          term == 'reproM' ~ 'Sex:Male',
                          term == 'reproPregnant' ~ "Pregnant")) %>%
  left_join(aic, by = "model") %>%
  mutate(dAIC = round(dAICc, 2),
         estimate = round(estimate, 2),
         std.error = round(std.error, 2)) %>%
  rename(predictor = term) %>%
  select(model, predictor, estimate, std.error, dAICc, df, weight)
  
# make a flex table
                          
m_table <- m_summary %>%
  flextable() %>%
  merge_v(j = c(1, 5, 6, 7)) %>%
  valign(j = c(1, 5, 6, 7), valign = "top") %>%
  fix_border_issues() %>%
  hline(i = 9) %>%
  vline(j = c(1, 4), part = "body") %>%
  width(j = 2, width = 2) %>%
  bold(i = 1, part = "header")

print(m_table, preview = "docx")


```











